---
title: "Untitled"
output: html_document
date: "2024-02-10"
---


```{r}
require(Matrix)
require(tidyverse)
require(Rspectra)
require(igraph)
require(ggrepel)
require(umap)
require(dplyr)
require(foreach)
library(RANN)

dir = "working_directory/prefix "

edges1x = read_csv(paste0(dir,"20clustsizemax no_readcount_z dist_strand gxgpairwise.csv"))

head(edges1x)
```
```{r}
edges1x2 = edges1x %>% rename(geneB = geneA, geneA = geneB)
head(edges1x2)

nrow(edges1x)*2
edges2x = rbind(edges1x, edges1x2) %>% distinct()
nrow(edges2x)

rm(edges1x2)
head(edges2x)
```
```{r}
bygeneA = edges2x %>% group_by(geneA) %>% summarise(maxz = max(zscore), 
                                                    coloc_partners = sum(pearson_resid > 1),
                                                    total_ints = sum(interactions), 
                                                    maxPR = max(pearson_resid))

head(bygeneA)
```

Louvain community analysis forces every gene into a hub. I recommend filtering for genes that have evidence of real colocalization specificity by filtering for max z-score or by number of colocalization partners. 
z-score reflects confidence of colocalization and is correlated to abundance. 
The Pearson residual is the deviation from expected colocalization and is not strongly correlated to abundance in our datasets. 
```{r}
bygeneA_filt = bygeneA %>% filter(maxz > 3.2)
nrow(bygeneA_filt)
nrow(bygeneA)

genes = bygeneA_filt$geneA

edges_filt = edges2x %>% filter(geneA %in% genes, geneB %in% genes)
head(edges_filt)
```

```{r}
PC_wider = edges_filt %>% select(geneA, geneB, pearson_resid) %>% pivot_wider(id_cols = geneA, names_from = geneB, values_from = pearson_resid) 
    
    #Get adjaceny matrix, genes by genes
    #turning into a matrix
    genexgene_mat = as.matrix(PC_wider[,2:ncol(PC_wider)])
    rownames(genexgene_mat) = PC_wider$geneA
    
    #making sure it's symmetric
    genexgene_mat = genexgene_mat[,rownames(genexgene_mat)]
    
    #don't want to be looking at deviation from expectation for non-colocalizing
    genexgene_mat[(genexgene_mat < 0)] <- 0
    
    genexgene_mat[is.na(genexgene_mat)] <- 0
    genexgene_mat[is.infinite(genexgene_mat)] <- 0
    
    dim(genexgene_mat)
```

saveRDS(genexgene_mat, file = "genexgene_mat.rds")


```{r}
total_variance = sum(genexgene_mat * genexgene_mat)

#PCA, 5 PCs
SVD = RSpectra::svds(genexgene_mat, 5)
  
#proportion of variance per PC
pve = SVD$d^2 / total_variance
print(plot(pve))

#cumulative variance explained
print(plot(cumsum(pve)))

#scaling with variance explained
svds_scaled = SVD$u * SVD$d
```
 

```{r}
PCA = svds_scaled

# Calculate the nearest neighbors  
nn_graph <- RANN::nn2(PCA, k = 100)  # k represents the number of nearest neighbors

# Extract the indices of the nearest neighbors
  indices <- nn_graph$nn.idx
  
  # Create an adjacency matrix for the nearest neighbor graph, neighbors = 1, everything else 0
  adj_matrix <- matrix(0, nrow = nrow(PCA), ncol = nrow(PCA))
  for (i in 1:nrow(PCA)) {
    adj_matrix[i, indices[i, ]] <- 1
  }
  
  #weighted by pearson_resid
  graph_object <- graph_from_adjacency_matrix(adj_matrix, mode = "undirected", weighted = T)
  
  #change resolution to change the number of hubs
  comms = cluster_louvain(graph_object, resolution = 0.5)
  sizes(comms)
```
  
```{r}
#making a UMAP
#other UMAP metrics are manhattan, euclidean, or pearson
UMAP <- umap(svds_scaled, metric = "cosine", n_neighbors = 50, input = "data")
```


```{r, fig.width = 5, fig.height = 5}
to_plot2 <- UMAP$layout %>% as_tibble() %>% mutate(geneA = rownames(genexgene_mat))
to_plot2$Community = as.factor(to_plot2$Community)

ggplot(to_plot2, aes(V1, V2, color = Community, fill = Community)) + 
  geom_point(alpha = 0.2) +  
  guides(colour = guide_legend(override.aes = list(alpha = 1))) + 
  theme_classic() + 
  #theme(legend.position="none") + 
  theme(text = element_text(size = 20, color = "black"))
```

write_csv(to_plot2, "UMAP coordinates.csv")

